---
title: "Final Assignment"
author: 'Yancheng Zhuang'
date: "22 March 2021"
output: pdf_document
fontsize: 11pt
highlight: tango
header-includes: \usepackage{bbm}
---

`r options(digits=3)`

## Exercise 1 
<!--
To investigate the effect of 3 types of diet, 78 persons were divided randomly in 3 groups, 
the first group following diet 1, second group diet 2 and the third group diet 3. 
Next to some other characteristics, the weight was measured before diet and after 6 weeks of diet for each person in the study. 
The collected data is summarized in the data frame diet.txt with the following columns: 
person – participant number, 
gender – gender (1 = male, 0 = female), 
age – age (years), 
height – height (cm), 
diet – the type of diet followed,
preweight – weight before the diet (kg), 
weight6weeks – weight after 6 weeks of diet (kg).
-->

**a)** 
<!--
By using only the columns preweight and weight6weeks, test the claim that the diet affects the weight loss. 
Give and check the assumptions of the test(s) applied.
-->

Firstly, the normality of the weight change is checked using QQ-plot. 

```{r, fig.height=4.5, fig.width=4.5, fig.align='center'}
diet <- read.table("diet.txt", header=TRUE)
diet$diet <- as.factor(diet$diet)
diet$gender <- as.factor(diet$gender)
with(diet, qqnorm(preweight - weight6weeks))
```

It can be seen that almost all points are approximately on a single line, so the normality can be assumed. 
Thus, it is appropriate to use paired t-test here. 
The null hypothesis is set up as the mean of the differences of the weight is 0. 

```{r}
res <- with(diet, t.test(preweight, weight6weeks, paired=TRUE)); res
```

Here it can be seen that the p-value = `r res$p.value` is smaller than 0.05, 
so the null hypothesis is rejected and I support the claim that the diet affects the weight loss. 

**b)** 
<!--
Compute and add to the data frame the variable weight.lost expressing the lost weight. 
Test the claim that the median lost weight is bigger than 3 kg. 
-->

The null hypothesis is set up as the median of this weight.lost is bigger than 3. 
Here the sign test is used to test the null hypothesis. 

```{r}
diet$weight.lost <- diet$preweight - diet$weight6weeks
res <- binom.test(
    sum(diet$weight.lost > 3), 
    length(diet$weight.lost), 
    alternative="less"
); res
```

According to the result, p = `r res$p.value` > 0.05. 
So, the null hypothesis cannot be rejected. 
The median lost weight is bigger than 3 kg. 

**c)** 
<!--
Using now weight.lost as response variable, 
apply one-way ANOVA to test whether type of diet has an effect on the lost weight. 
Do all three types diets lead to weight loss?
Which diet was the best for losing weight?
-->

The null hypothesis is set up as the type of diet has no effect on the lost weight.

```{r}
diet_lm <- lm(weight.lost ~ diet, data=diet)
res <- anova(diet_lm); res
```

The p-value = `r res["diet", "Pr(>F)"]` < 0.05, which leads to rejecting the null hypothesis. 
Factor diet is significant. 

```{r}
diet_ <- diet; contrasts(diet_$diet) <- contr.sum
diet_lm <- lm(weight.lost ~ diet, data=diet_)
res <- summary(diet_lm)$coefficients; res
```

According to the results, both diet 1 and diet 2 don't lead to a weight loss. 
diet 3 is the best for losing weight. 

The normality of the residuals should be checked. 

```{r, fig.height=3, fig.width=6, fig.align='center'}
par(mfrow=c(1, 2))
qqnorm(residuals(diet_lm)); plot(fitted(diet_lm), residuals(diet_lm))
```

It can be seen from the left plot that almost all points are approximately on a single line, so the normality can be assumed.
The right plot also seems okay. 

**d)** 
<!--
Use two-way ANOVA to investiage effect of the diet and gender (and possible interaction) on the lost weight. 
Is the Friedman test relevant here? Can it be at all applied? 
-->

The null hypothesis is set up as: 

- $H_A$: no main effect of the first factor diet

- $H_B$: no main effect of the second factor gender

- $H_{AB}$: no interaction between the two factors diet and gender

```{r}
diet_lm <- lm(weight.lost ~ diet * gender, data=diet)
res <- anova(diet_lm); res
```

According to results, p-value for $H_A$ is `r res["diet", "Pr(>F)"]` < 0.05, 
so $H_A$ is rejected. Factor diet has significant effect on weight loss. 
p-value for $H_B$ is `r res["gender", "Pr(>F)"]` > 0.05, 
so $H_B$ is not rejected. Factor gender has no main effect on weight loss. 
p-value for $H_AB$ is `r res["diet:gender", "Pr(>F)"]` < 0.05, 
so $H_AB$ is rejected. The interaction between diet and gender has effect on weight loss. 

The normality of the residuals should be checked. 

```{r fig.height=3, fig.width=6, fig.align='center'}
par(mfrow=c(1, 2))
qqnorm(residuals(diet_lm)); plot(fitted(diet_lm), residuals(diet_lm))
```

In the left plot, some points are stray away from the line, but most of the points seem ok. 
The right plot seems good, the spread in the residuals does not change systematically. 

Friedman test is not relevant here. 
The aim of Friedman test is to understand the dependence of weight lost on the treatment factor 
with block variable used to create homogeneous groups of experimental units. 
Here is to investiage effect of the diet and gender (and their interaction) on the weight lost. 
So it's not relevent. 
Also, Friedman test does not take interaction between factors into account. 
However, here there does exist interaction between diet and gender. 
So, it can not be applied. 

**e)** 
<!--
Apply appropriate model to investigate effect(s) of diet and height (and possibly their interaction) on the lost weight. 
Is the effect of height the same for all 3 types of diet?
-->

The null hypothesis is set up as: 

- $H_A$: no main effect of the first factor diet

- $H_B$: no main effect of the second factor height

- $H_{AB}$: no interaction between the two factors diet and height

```{r}
res <- anova(lm(weight.lost ~ diet * height , data=diet)); res
```

According to results, p-value for $H_A$ is `r res["diet", "Pr(>F)"]` < 0.05, 
so $H_A$ is rejected. Factor diet has significant effect on weight loss. 
p-value for $H_B$ is `r res["height", "Pr(>F)"]` > 0.05, 
so $H_B$ is not rejected. Factor height has no main effect on weight loss. 
p-value for $H_AB$ is `r res["diet:height", "Pr(>F)"]` > 0.05, 
so $H_AB$ is not rejected. There is no interaction between diet and height. 

So we again use the addition model. 

```{r}
diet_lm <- lm(weight.lost ~ height + diet, data=diet)
res <- anova(diet_lm); res
```

The results are same as above. 
Only $H_A$ is rejected. 

```{r, fig.height=3, fig.width=6, fig.align='center'}
par(mfrow=c(1, 2))
with(diet, plot(weight.lost ~ height, pch=unclass(diet)))
for (i in 1:2) abline(lm(weight.lost~height, data=diet[diet$diet==i, ]))

with(diet, plot(weight.lost ~ height, pch=unclass(diet)))
for (i in 1:3) abline(lm(weight.lost~height, data=diet[diet$diet==i, ]))
```

According to the plot, the effect of height is not the same for all 3 types of diet. 
The effect of height is similar while using diet1 and diet2 which is shown in the left plot, 
but it is diffenet for diet: the slope becomes negative (in the right plot). 

Lastly, it is still necessary to check the normality. 

```{r fig.height=3, fig.width=6, fig.align='center'}
par(mfrow=c(1, 2))
qqnorm(residuals(diet_lm)); plot(fitted(diet_lm), residuals(diet_lm))
```

Both plots seem good. 

**f)**
<!--
Which of the two approaches, the one from c) or from e), do you prefer? Why? 
For the preferred model, predict the lost weight for all three types of diet for an average person.
-->

```{r}
summary(lm(weight.lost ~ diet, data=diet))$adj.r.squared
```

```{r}
summary(lm(weight.lost ~ height + diet, data=diet))$adj.r.squared
```

I perfer the approach in e) but i will adopt the model in c). 
The approach in e) considered more variables and investigated the potential interaction between diet and height compared to c), 
but they achieves the results that only diet variable has effect on the lost weight. 
So model in c) in perferred as it has less variables and the ajusted R squared is a little bit higher. 

```{r}
coef <- summary(lm(weight.lost ~ diet, data=diet))$coefficients; coef
```

So for an average person, for diet 1, the perdicted lose weight is `r coef["(Intercept)", "Estimate"]`; 
for diet 2, the perdicted lose weight is `r coef["(Intercept)", "Estimate"] + coef["diet2", "Estimate"]`; 
for diet 3, the perdicted lose weight is `r coef["(Intercept)", "Estimate"] + coef["diet3", "Estimate"]`. 

**g)**
<!--
Create the categorical (binary) variable lost.4kg expressing whether the lost weight is bigger than 4kg or not. 
(Hint: if a=c(1,-1,2), then a>0 = TRUE FALSE TRUE.)
Suppose now that we observed the variable lost.4kg instead of weight6weeks. 
Could we still test the hypothesis in b), c), d), e)? 
If yes, propose relevant models and perform relevant tests.
--> 

```{r}
diet$lost.4kg <- as.factor(diet$weight.lost > 4)
```

For b), the original null hypothesis is set up as the median of this weight.lost is bigger than 3. 
But here we only have lost.4kg that counts the number of people whose weight lost is bigger than 4. 
So, the original null hypothesis can not be tested. 
We can only test null hypothesis like median of lost weight is bigger than 4. 

```{r}
res <- binom.test(
    sum(diet$lost.4kg==TRUE), 
    length(diet$lost.4kg), 
    alternative="less"
); res
```

According to the result, p = `r res$p.value` > 0.05. 
So, the null hypothesis cannot be rejected. 
The median lost weight is bigger than 4 kg. 
So, the median lost weight is bigger than 3 kg. 

For c), d), e), the null hypothesis are whether the diet, gender or height have effects on the lost weight. 
Because here we use binary lost.4kg as the response variable, we can make use logistic regression to 
investigate the effect of diet, gender or height variable. 

The null hypothesis is set up as: 

- $H_A$: no main effect of the first factor diet

- $H_B$: no main effect of the second factor gender

- $H_C$: no main effect of the second factor height

- $H_{AB}$: no interaction between the two factors diet and gender

- $H_{AC}$: no interaction between the two factors diet and height

```{r}
diet_glm <- glm(
    lost.4kg ~ diet + gender + height + diet:gender + diet:height, 
    data=diet, family=binomial
)
anova(diet_glm, test="Chisq")
```

According to the deviance table, $H_A$ and $H_{AB}$ are rejected, 
since their p values are smaller than 0.05. 
Other hypothesis are not rejected. 
So there is no interaction between diet and height. 
gender and height has no effect on whether the lost weight is bigger than 4kg or not. 
